Fast Sorting on Modern Computers

### 2005-08-30 docs/tech_report/heap.tex
spell check

Meanwhile, Figure \ref{memory-tuned heapsort curve} shows us the rate of correct
predictions, and each point is calculated using a seperate branch predictor. As
a result, we expect $1-y = \frac{1}{x+1}$, where $x$ and $y$ are the cartesian
coordinates of each point on the graph.  $1-y$ is the number of mispredictions
from the graph, while $\frac{1}{x+1}$ is the increase in $H_N$ for step $N$.

When the array is short, as in this example, the branch predictor has too much
feedback to be close to $H_N$. For $N=2$, we would approximate a 33\%
misprediction ratio using $H_N$, but we measure 39\%. The approximated and
actual ratios become closer as $N$ increases, as we would expect, and for $N=7$,
we approximate 12.5\% and measure 12.7\%. We conclude that initial suggestion is
correct: while the branch prediction ratio of finding the minimum is very close
to $H_N-1$, the relationship is less accurate for very short arrays.

@@

The formula we use to explain selection sort's branch prediction preformance
breaks down for short arrays, as in our 8-heap. It should be possible to more
accurately model and therefore predict the branch prediction ratio when the
array is short. Our results show us what we expect the answer to be, but we have
not attempted to create a formula to predict it.

>>>>

Meanwhile, Figure \ref{memory-tuned heapsort curve} shows us the rate of correct
predictions, and each point is calculated using a separate branch predictor. As
a result, we expect $1-y = \frac{1}{x+1}$, where $x$ and $y$ are the Cartesian
coordinates of each point on the graph.  $1-y$ is the number of mispredictions
from the graph, while $\frac{1}{x+1}$ is the increase in $H_N$ for step $N$.

When the array is short, as in this example, the branch predictor has too much
feedback to be close to $H_N-1$. For $N=2$, we would approximate a 33\%
misprediction ratio using $H_N-1$, but we measure 39\%. The approximated and
actual ratios become closer as $N$ increases, as we would expect, and for $N=7$,
we approximate 12.5\% and measure 12.7\%. We conclude that initial suggestion is
correct: while the branch prediction ratio of finding the minimum is very close
to $H_N-1$, the relationship is less accurate for very short arrays.

@@

The formula we use to explain selection sort's branch prediction performance
breaks down for short arrays, as in our 8-heap. It should be possible to more
accurately model and therefore predict the branch prediction ratio when the
array is short. Our results show us what we expect the answer to be, but we have
not attempted to create a formula to predict it.
